Reproducible Research: Assignment 1
==========================================

```
library(ggplot2)
library(scales)
library(Hmisc)
```

## Required Settings
```{r}
echo = TRUE  
options(scipen = 1)  
```

## Load & Process Data 
```{r}
unzip("activity.zip")
dataTable <- read.csv("activity.csv", colClasses = c("integer", "Date", "factor"))
dataTable$month <- as.numeric(format(dataTable$date, "%m"))
noNA <- na.omit(dataTable)
rownames(noNA) <- 1:nrow(noNA)
head(noNA)
dim(noNA)
library(ggplot2)
```


### What is mean total number of steps taken per day?

* Histogram of the total number of steps taken each day
```{r}
ggplot(noNA, aes(date, steps)) + 
geom_bar(stat = "identity", colour = "steelblue", fill = "steelblue", width = 0.7) + 
facet_grid(. ~ month, scales = "free") + 
labs(title = "Histogram of Total Number of Steps Taken Each Day", x = "Date", y = "Total Number of Steps")
```

* Calculate and report the mean and median total number of steps taken per day

Mean total number of steps taken per day:
```{r}
totalSteps <- aggregate(noNA$steps, list(Date = noNA$date), FUN = "sum")$x
mean(totalSteps)
```
Median total number of steps taken per day:
```{r}
median(totalSteps)
```

### What is the average daily activity pattern?

```{r}
avgSteps <- aggregate(noNA$steps, list(interval = as.numeric(as.character(noNA$interval))), FUN = "mean")
names(avgSteps)[2] <- "meanOfSteps"

ggplot(avgSteps, aes(interval, meanOfSteps)) + 
geom_line(color = "steelblue", size = 0.8) + 
labs(title = "Time Series Plot of the 5-minute Interval", x = "5-minute intervals", y = "Average Number of Steps Taken")
```

* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
avgSteps[avgSteps$meanOfSteps == max(avgSteps$meanOfSteps), ]
```

### Imputing missing values

```{r}
sum(is.na(dataTable))
```
```{r}
newDataTable <- dataTable
for (i in 1:nrow(newDataTable)) {
    if (is.na(newDataTable$steps[i])) {
        newDataTable$steps[i] <- avgSteps[which(newDataTable$interval[i] == avgSteps$interval), ]$meanOfSteps
    }
}

head(newDataTable)
sum(is.na(newDataTable))
```

```{r}
ggplot(newDataTable, aes(date, steps)) + 
geom_bar(stat = "identity",colour = "steelblue", fill = "steelblue",width = 0.7) + 
facet_grid(. ~ month, scales = "free") + 
labs(title = "Histogram of Total Number of Steps Taken Each Day (without missing data)", x = "Date", y = "Total Number of Steps")
```

Mean total number of steps taken per day:
```{r}
newTotalSteps <- aggregate(newDataTable$steps, 
                           list(Date = newDataTable$date), 
                           FUN = "sum")$x
newMean <- mean(newTotalSteps)
newMean
```
Median total number of steps taken per day:
```{r}
newMedian <- median(newTotalSteps)
newMedian
```
Compare them with the two before imputing missing data:
```{r}
oldMean <- mean(totalSteps)
oldMedian <- median(totalSteps)
newMean - oldMean
newMedian - oldMedian
```

### Are there differences in activity patterns between weekdays and weekends?

```{r}
head(newDataTable)
newDataTable$weekdays <- factor(format(newDataTable$date, "%A"))
levels(newDataTable$weekdays)
levels(newDataTable$weekdays) <- list(weekday = c("Monday", "Tuesday",
                                             "Wednesday", 
                                             "Thursday", "Friday"),
                                 weekend = c("Saturday", "Sunday"))
levels(newDataTable$weekdays)
table(newDataTable$weekdays)
```

```{r}
avgSteps <- aggregate(newDataTable$steps, 
                      list(interval = as.numeric(as.character(newDataTable$interval)), 
                           weekdays = newDataTable$weekdays),
                      FUN = "mean")
names(avgSteps)[3] <- "meanOfSteps"
library(lattice)
xyplot(avgSteps$meanOfSteps ~ avgSteps$interval | avgSteps$weekdays, 
       layout = c(1, 2), type = "l", 
       xlab = "Interval", ylab = "Number of steps")
```
